```{r}
library(readxl)
scores <-  read.delim("aggregated_scores.txt")
farm <-  read_excel("2026_02_04_–¥–ª—è_–ì–µ–Ω–æ—Ç–µ–∫–∞.xlsx")


```

```{r}
library(dplyr)
farm <- farm %>% rename('IID' = 'ID')
scores$IID <- sub("_.*", "", scores$IID)
```

```{r}
merged <- merge(farm, scores, by = "IID")
# convert the "treatment_program_completion_fact" column to binary format
merged$outcome <- ifelse(merged$`—Ñ–∞–∫—Ç_–∑–∞–≤–µ—Ä—à–µ–Ω–∏—è –ø—Ä–æ–≥—Ä–∞–º–º—ã –ª–µ—á–µ–Ω–∏—è` == "–¥–∞", 1, 0)
merged <- merged %>% rename('drug' = '–≥—Ä—É–ø–ø–∞ (–ü–Ý–ï–ü–ê–Ý–ê–¢, –ö–û–¢–û–Ý–´–ô –ü–û–õ–£–ß–ê–õ –ü–ê–¶–ò–ï–ù–¢)')  
merged$drug <- as.factor(merged$drug)
```

Statistics - how many people took which drug and how many of them completed treatment
```{r}
library(dplyr)
library(ggplot2)

plot_df <- farm %>%
    mutate(
        treatment = recode(
            `–≥—Ä—É–ø–ø–∞ (–ü–Ý–ï–ü–ê–Ý–ê–¢, –ö–û–¢–û–Ý–´–ô –ü–û–õ–£–ß–ê–õ –ü–ê–¶–ò–ï–ù–¢)`,
            "–¶–∏–∞–Ω–∞–º–∏–¥" = "Cyanamide",
            "–ü—Ä–µ–≥–∞–±–∞–ª–∏–Ω (–õ–∏—Ä–∏–∫–∞)" = "Pregabalin(Lyrica)",
            "–î–∏—Å—É–ª—å—Ñ–∏—Ä–∞–º" = "Disulfiram",
            "–ü–ª–∞—Ü–µ–±–æ" = "Plasebo"
        ),
        completed = recode(
            `—Ñ–∞–∫—Ç_–∑–∞–≤–µ—Ä—à–µ–Ω–∏—è –ø—Ä–æ–≥—Ä–∞–º–º—ã –ª–µ—á–µ–Ω–∏—è`,
            "–¥–∞" = "Yes",
            "–Ω–µ—Ç" = "No"
        )
    ) %>%
    count(treatment, completed)

ggplot(
    plot_df,
    aes(
        x = treatment,
        y = n,
        fill = completed
    )
) +
    geom_col(color = "grey40", linewidth = 0.3) +
    geom_text(
        aes(label = n),
        position = position_stack(vjust = 0.5),
        size = 5,
        fontface = "bold"
    ) +
    scale_fill_manual(
        values = c(
            "Yes" = "#BFD8B8",  
            "No"  = "#F4C2C2"   
        )
    ) +
    labs(
        x = "Treatment",
        y = "Number of Patients",
        fill = "Completed Treatment"
    ) +
    theme_minimal(base_size = 16) +
    theme(
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 14),
        axis.text.x = element_text(size = 14),
        panel.grid.major.x = element_blank()
    )
```



Linear recursion for all PRS
predict the outcome (outcomes 0 or 1) depending on the PRS score
```{r}
library(pROC)
library(dplyr)

# –ü—É—Å—Ç–∞—è —Ç–∞–±–ª–∏—Ü–∞
results_table <- data.frame(
    Score = character(),
    OR = numeric(),
    OR_lower = numeric(),
    OR_upper = numeric(),
    p_value = numeric(),
    AUC = numeric(),
    AUC_lower = numeric(),
    AUC_upper = numeric(),
    N = numeric(),
    stringsAsFactors = FALSE
)

pgs_list <- unique(merged$PGS)

for (pgs in pgs_list) {
    
    
    df <- merged %>% filter(PGS == pgs)
    
  
    if(all(is.na(df$SUM)) || length(unique(df$SUM)) < 2) next
    
    
    df$PRS_z <- scale(df$SUM)
    
  
    model <- glm(outcome ~ PRS_z, family = binomial(), data = df)
    
    # OR
    coef_val <- coef(model)[2]
    OR <- exp(coef_val)
    
    # p-value
    coef_table <- coef(summary(model))
    p_val <- coef_table[2, "Pr(>|z|)"]
    
    # CI
    conf <- tryCatch({
        exp(confint(model)[2, ])
    }, error = function(e) c(NA, NA))
    
    # ROC
    df$pred <- predict(model, type = "response")
    roc_obj <- roc(df$outcome, df$pred)
    
    auc_val <- auc(roc_obj)
    auc_ci <- ci.auc(roc_obj)
    
    N <- nrow(df)
    
 
    results_table <- rbind(results_table, data.frame(
        Score = pgs,
        OR = OR,
        OR_lower = conf[1],
        OR_upper = conf[2],
        p_value = p_val,
        AUC = as.numeric(auc_val),
        AUC_lower = auc_ci[1],
        AUC_upper = auc_ci[3],
        N = N
    ))
}

# –í—ã–≤–æ–¥
print(results_table)
```




+ covariate - column drug 
+ Benjamin Hochberg's correction for multiple comparisons
 
```{r}
library(pROC)
library(dplyr)

results_table <- data.frame(
    Score = character(),
    OR = numeric(),
    OR_lower = numeric(),
    OR_upper = numeric(),
    p_value = numeric(),
    AUC = numeric(),
    AUC_lower = numeric(),
    AUC_upper = numeric(),
    N = numeric(),
    stringsAsFactors = FALSE
)

pgs_list <- unique(merged$PGS)

for (pgs in pgs_list) {
    
    df <- merged %>% filter(PGS == pgs)
    
    if(all(is.na(df$SUM)) || length(unique(df$SUM)) < 2) next
    

    df$PRS_z <- scale(df$SUM)
    
    #logistic regression with covariate drug
    model <- glm(outcome ~ PRS_z + drug, family = binomial(), data = df)
    
    coef_val <- coef(model)["PRS_z"]
    OR <- exp(coef_val)
    
    coef_table <- coef(summary(model))
    p_val <- coef_table["PRS_z", "Pr(>|z|)"]
    
    conf <- tryCatch({
        exp(confint(model)["PRS_z", ])
    }, error = function(e) c(NA, NA))
    
    # ROC
    df$pred <- predict(model, type = "response")
    roc_obj <- roc(df$outcome, df$pred)
    
    auc_val <- auc(roc_obj)
    auc_ci <- ci.auc(roc_obj)
    
    N <- nrow(df)
    
    results_table <- rbind(results_table, data.frame(
        Score = pgs,
        OR = OR,
        OR_lower = conf[1],
        OR_upper = conf[2],
        p_value = p_val,
        AUC = as.numeric(auc_val),
        AUC_lower = auc_ci[1],
        AUC_upper = auc_ci[3],
        N = N
    ))
}

# Benjamini-Hochberg amendment
results_table$p_adjusted <- p.adjust(results_table$p_value, method = "BH")

# sort by importance
results_table <- results_table[order(results_table$p_adjusted), ]


print(results_table)
```
```{r}
library(dplyr)
library(pROC)


# summary table (PRS level)

results_table <- data.frame(
    Score = character(),
    OR_PRS = numeric(),
    CI_lower = numeric(),
    CI_upper = numeric(),
    p_PRS = numeric(),
    p_interaction_global = numeric(),
    p_interaction_min = numeric(),
    AUC = numeric(),
    AUC_lower = numeric(),
    AUC_upper = numeric(),
    N = numeric(),
    stringsAsFactors = FALSE
)

# ---------------------------
# table interaction for drugs 
# ---------------------------
interaction_table_all <- data.frame(
    Score = character(),
    term = character(),
    p_value = numeric(),
    stringsAsFactors = FALSE
)

pgs_list <- unique(merged$PGS)

for (pgs in pgs_list) {
    
    df <- merged %>% filter(PGS == pgs)
    
   
    if (all(is.na(df$SUM)) || length(unique(df$SUM)) < 2) next
    
    # standartisation PRS
    df$PRS_z <- scale(df$SUM)
    
    # factor + reference = Placebo
    df$drug <- as.factor(df$drug)
    df$drug <- relevel(df$drug, ref = "–ü–ª–∞—Ü–µ–±–æ")
    
    # ---------------------------
    # –º–æ–¥–µ–ª—å —Å interaction
    # ---------------------------
    model <- glm(outcome ~ PRS_z * drug,
                 family = binomial(),
                 data = df)
    
    coef_table <- summary(model)$coefficients
    
    # ---------------------------
    # PRS effect (in placebo)
    # ---------------------------
    OR_PRS <- exp(coef_table["PRS_z", "Estimate"])
    p_PRS <- coef_table["PRS_z", "Pr(>|z|)"]
    
    # CI –¥–ª—è PRS
    CI <- tryCatch({
        exp(confint(model)["PRS_z", ])
    }, error = function(e) c(NA, NA))
    
    # ---------------------------
    # interaction effects (for each drug)
    # ---------------------------
    interaction_rows <- grep("PRS_z:drug", rownames(coef_table), value = TRUE)
    
    if (length(interaction_rows) > 0) {
        
        interaction_pvals <- coef_table[interaction_rows, "Pr(>|z|)"]
        
        interaction_tmp <- data.frame(
            Score = pgs,
            term = interaction_rows,
            p_value = interaction_pvals
        )
        
        interaction_table_all <- rbind(interaction_table_all, interaction_tmp)
        
        p_interaction_min <- min(interaction_pvals, na.rm = TRUE)
        
    } else {
        p_interaction_min <- NA
    }
    
    # global test interaction (LRT)
    model_reduced <- glm(outcome ~ PRS_z + drug,
                         family = binomial(),
                         data = df)
    
    p_interaction_global <- anova(model_reduced, model, test = "LRT")[2, "Pr(>Chi)"]
    

    # ROC / AUC
   
    df$pred <- predict(model, type = "response")
    roc_obj <- roc(df$outcome, df$pred)
    
    auc_val <- auc(roc_obj)
    auc_ci <- ci.auc(roc_obj)
    

    results_table <- rbind(results_table, data.frame(
        Score = pgs,
        OR_PRS = OR_PRS,
        CI_lower = CI[1],
        CI_upper = CI[2],
        p_PRS = p_PRS,
        p_interaction_global = p_interaction_global,
        p_interaction_min = p_interaction_min,
        AUC = as.numeric(auc_val),
        AUC_lower = auc_ci[1],
        AUC_upper = auc_ci[3],
        N = nrow(df)
    ))
}


# correction BH

results_table$p_PRS_adj <- p.adjust(results_table$p_PRS, method = "BH")
results_table$p_interaction_adj <- p.adjust(results_table$p_interaction_global, method = "BH")

interaction_table_all$p_adj <- p.adjust(interaction_table_all$p_value, method = "BH")


results_table <- results_table %>%
    arrange(p_interaction_adj, p_PRS_adj)

interaction_table_all <- interaction_table_all %>%
    arrange(p_value)


print(results_table)
print(interaction_table_all)
```

Forst Plot - I'm not sure it's needed, it's a bit pointless
```{r}
library(ggplot2)
library(dplyr)


plot_df <- results_table %>%
    mutate(
        OR_low = CI_lower,
        OR_high = CI_upper
    ) %>%
    arrange(OR_PRS) %>%
    mutate(Score = factor(Score, levels = Score))

# forest plot
ggplot(plot_df, aes(x = OR_PRS, y = Score)) +
    geom_point() +
    geom_errorbarh(aes(xmin = OR_low, xmax = OR_high), height = 0.2) +
    geom_vline(xintercept = 1, linetype = "dashed") +
    xlab("Odds Ratio (PRS)") +
    ylab("Polygenic Score") +
    theme_bw()
```
only for  prs PGS002752_hmPOS_GRCh37 PGS000204_hmPOS_GRCh37 PGS003376_hmPOS_GRCh37

```{r}
library(dplyr)
library(pROC)


pgs_list <- c(
    "PGS002152_hmPOS_GRCh37",
    "PGS002752_hmPOS_GRCh37",
    "PGS000204_hmPOS_GRCh37",
    "PGS003376_hmPOS_GRCh37"
)

drug_list <- unique(merged$drug)


results_table <- data.frame(
    Drug = character(),
    Score = character(),
    OR = numeric(),
    CI_lower = numeric(),
    CI_upper = numeric(),
    p_value = numeric(),
    AUC = numeric(),
    AUC_lower = numeric(),
    AUC_upper = numeric(),
    N = numeric(),
    stringsAsFactors = FALSE
)

cairo_pdf("ROC_by_drug.pdf", width = 7, height = 7)



for (drug_name in drug_list) {
    
    drug_df <- merged %>%
        filter(drug == drug_name)
    
    tmp_results <- data.frame()
    
    roc_list <- list()
    auc_values <- c()
    labels <- c()
    colors <- c("red", "blue", "darkgreen", "black")
    
    i <- 1

  

    for (pgs in pgs_list) {
        
        df <- drug_df %>%
            filter(PGS == pgs)
        
        if (nrow(df) < 10 || all(is.na(df$SUM)) || length(unique(df$outcome)) < 2) next
        
        df$PRS_z <- scale(df$SUM)
        
        model <- glm(outcome ~ PRS_z,
                     family = binomial(),
                     data = df)
        
        coef_table <- summary(model)$coefficients
        
        OR <- exp(coef_table["PRS_z", "Estimate"])
        p_val <- coef_table["PRS_z", "Pr(>|z|)"]
        
        CI <- tryCatch({
            exp(confint(model)["PRS_z", ])
        }, error = function(e) c(NA, NA))
        
        df$pred <- predict(model, type = "response")
        roc_obj <- roc(df$outcome, df$pred)
        
        roc_list[[i]] <- roc_obj
        auc_values[i] <- auc(roc_obj)
        labels[i] <- sub("_.*", "", pgs)
        
        auc_ci <- ci.auc(roc_obj)
        
        tmp_results <- rbind(tmp_results,
                             data.frame(
                                 Drug = drug_name,
                                 Score = pgs,
                                 OR = OR,
                                 CI_lower = CI[1],
                                 CI_upper = CI[2],
                                 p_value = p_val,
                                 AUC = as.numeric(auc(roc_obj)),
                                 AUC_lower = auc_ci[1],
                                 AUC_upper = auc_ci[3],
                                 N = nrow(df)
                             ))
        
        i <- i + 1
    }
    
    # ---------------------------------
    # BH correction inside drug
    # ---------------------------------
    if (nrow(tmp_results) > 0) {
        tmp_results$p_adj_BH <- p.adjust(tmp_results$p_value, method = "BH")
        results_table <- rbind(results_table, tmp_results)
    }
    
    # ---------------------------------
    # ROC PLOT 
    # ---------------------------------
    if (length(roc_list) > 0) {
        
        plot(smooth(roc_list[[1]]),
             col = colors[1],
             lwd = 4,
             main = paste("ROC curves -", drug_name),
             legacy.axes = TRUE,
             xlab = "False positive rate",
             ylab = "True positive rate",
             cex.main = 1.8, 
             cex.lab  = 1.6,   
             cex.axis = 1.4 )
        
        if (length(roc_list) > 1) {
            for (j in 2:length(roc_list)) {
                plot(smooth(roc_list[[j]]),
                     col = colors[j],
                     lwd = 3,
                     add = TRUE)
            }
        }
        
        #abline(a = 0, b = 1, lty = 2, col = "gray",lwd = 2)
        
        legend("bottomright",
               legend = paste0(labels),
               col = colors,
               lwd = 4,
               cex = 1.3,
               bty = "n")
    }
}

dev.off()


results_table <- results_table %>%
    arrange(Drug, p_adj_BH)

print(results_table)
```


