
```{r}
alco_data <- read.csv("C:/Users/Ast/Desktop/АЛКОГОЛИКИ/data_prs/alcofinal_summary.csv")
```

```{r}
merged_data <- read.delim("C:/Users/Ast/Desktop/АЛКОГОЛИКИ/data_prs/merged_pheno_prs.txt", header=T)
```

```{r}
merged_data$audit_score <- rowSums(merged_data[, 2:11], na.rm = TRUE)
merged_data$alc_binary <- ifelse(merged_data$audit_score >= 15, 1, 0)
merged_data$gender_binary <- ifelse(merged_data$gender == "female", 0, 1)
```

```{r}
merged_data$age <- as.integer(
  floor(
    as.numeric(
      difftime(
        as.Date(merged_data$alcoholConsumption_updateDate),
        as.Date(merged_data$birthdate),
        units = "days"
      )
    ) / 365.25
  )
)
```


```{r}
library(pROC)
library(dplyr)

# Создаем пустую таблицу для итогов
results_table <- data.frame(
    Score = character(),
    OR = numeric(),
    OR_lower = numeric(),
    OR_upper = numeric(),
    p_value = numeric(),
    AUC = numeric(),
    AUC_lower = numeric(),
    AUC_upper = numeric(),
    N = numeric(),
    stringsAsFactors = FALSE
)


pdf("PRS_ROC_curves.pdf", width = 7, height = 7)

#  39 - 50 columns
for (col_idx in 39:50) {
    
    score_name <- colnames(merged_data)[col_idx]
    
    # Check empty 
    if(all(is.na(merged_data[[col_idx]])) || length(unique(merged_data[[col_idx]])) < 2) next
    
    #  Standartisation
    merged_data[[paste0(score_name, "_z")]] <- scale(merged_data[[col_idx]])
    
    #  Log regression
    formula_text <- paste("alc_binary ~", paste0(score_name, "_z"))
    model <- glm(as.formula(formula_text), family = binomial(), data = merged_data)
    
    #  OR и 95% CI
    coef_table <- coef(summary(model))
    coef_val <- coef(model)[2]                  
    OR <- exp(coef_val)
    
    # p-value
    p_val <- coef_table[2, "Pr(>|z|)"]
    
    # 95% CI с защитой от ошибок
    conf <- tryCatch({
        exp(confint(model)[2, ])
    }, error = function(e) c(NA, NA))
    
    # ROC AUC
    model_data <- model$model
    model_data$pred_prob <- predict(model, type = "response")
    roc_obj <- roc(model_data$alc_binary ~ model_data$pred_prob)
    auc_val <- auc(roc_obj)
    auc_ci <- ci.auc(roc_obj)
    
    N <- nrow(model_data)
    
    
    results_table <- rbind(results_table, data.frame(
        Score = score_name,
        OR = OR,
        OR_lower = conf[1],
        OR_upper = conf[2],
        p_value = p_val,
        AUC = as.numeric(auc_val),
        AUC_lower = auc_ci[1],
        AUC_upper = auc_ci[3],
        N = N
    ))
  
    #  ROC
    plot(roc_obj,
         col = "blue",
         lwd = 2,
         main = paste("ROC Curve for", score_name),
         print.auc = TRUE)
    abline(a = 0, b = 1, lty = 2, col = "red")
}


dev.off()


write.csv(results_table, "PRS_summary_table.csv", row.names = FALSE)


print(results_table)
```

With predictors - AGE GENDER
+ Benjamini-Hochberg correction
```{r}
library(pROC)
library(dplyr)

results_table <- data.frame(
    Score = character(),
    OR = numeric(),
    OR_lower = numeric(),
    OR_upper = numeric(),
    p_value = numeric(),
    AUC = numeric(),
    AUC_lower = numeric(),
    AUC_upper = numeric(),
    N = numeric(),
    stringsAsFactors = FALSE
)

pdf("PRS_ROC_curves_adjusted.pdf", width = 7, height = 7)

for (col_idx in 39:50) {
    
    score_name <- colnames(merged_data)[col_idx]
    
    
    if(all(is.na(merged_data[[col_idx]])) || length(unique(merged_data[[col_idx]])) < 2) next
    
   
    merged_data[[paste0(score_name, "_z")]] <- scale(merged_data[[col_idx]])
    
    # model with covariates
    formula_text <- paste("alc_binary ~", paste0(score_name, "_z"), "+ age + gender")
    
    model <- glm(as.formula(formula_text),
                 family = binomial(),
                 data = merged_data)
    
    coef_table <- coef(summary(model))
    
  
    coef_idx <- grep(paste0(score_name, "_z"), rownames(coef_table))
    
    # OR
    coef_val <- coef(model)[coef_idx]
    OR <- exp(coef_val)
    
    # p-value
    p_val <- coef_table[coef_idx, "Pr(>|z|)"]
    
    # CI
    conf <- tryCatch({
        exp(confint(model)[coef_idx, ])
    }, error = function(e) c(NA, NA))
    
    # ROC 
    model_data <- model$model
    model_data$pred_prob <- predict(model, type = "response")
    
    roc_obj <- roc(model_data$alc_binary, model_data$pred_prob)
    auc_val <- auc(roc_obj)
    auc_ci <- ci.auc(roc_obj)
    
    N <- nrow(model_data)
    

    results_table <- rbind(results_table, data.frame(
        Score = score_name,
        OR = OR,
        OR_lower = conf[1],
        OR_upper = conf[2],
        p_value = p_val,
        AUC = as.numeric(auc_val),
        AUC_lower = auc_ci[1],
        AUC_upper = auc_ci[3],
        N = N
    ))
    
    # ROC график
    plot(roc_obj,
         col = "blue",
         lwd = 2,
         main = paste("ROC Curve (adjusted) for", score_name),
         print.auc = TRUE)
    abline(a = 0, b = 1, lty = 2, col = "red")
}

dev.off()

# FDR (Benjamini-Hochberg correction)
results_table$p_adj_BH <- p.adjust(results_table$p_value, method = "BH")

print(results_table)
```
a single graph for six curves 
  "PGS000204",
  "PGS002752",
  "PGS003376"
with and without covariates
```{r}
library(pROC)
model_labels <- c(
  "PGS000204",
  "PGS002752",
  "PGS003376"
)

scores <- c("PGS000204_hmPOS_GRCh37_SUM",
            "PGS002752_hmPOS_GRCh37_SUM",
            "PGS003376_hmPOS_GRCh37_SUM")

roc_list <- list()
auc_values <- c()
labels <- c()

colors <- c("black", "gold", "green", "red", "orange", "blue")

i <- 1

for (score_name in scores) {
  
  merged_data[[paste0(score_name, "_z")]] <- scale(merged_data[[score_name]])
  
  # PRS only
  model1 <- glm(
    as.formula(paste("alc_binary ~", paste0(score_name, "_z"))),
    family = binomial(),
    data = merged_data
  )
  
  model_data1 <- model1$model
  model_data1$pred <- predict(model1, type = "response")
  
  roc1 <- roc(model_data1$alc_binary, model_data1$pred)
  
  roc_list[[i]] <- roc1
  auc_values[i] <- auc(roc1)
  labels[i] <-  paste0(model_labels[which(scores == score_name)], "")
  i <- i + 1
  
  # PRS + covariates
  model2 <- glm(
    as.formula(paste("alc_binary ~", paste0(score_name, "_z"), "+ age + gender")),
    family = binomial(),
    data = merged_data
  )
  
  model_data2 <- model2$model
  model_data2$pred <- predict(model2, type = "response")
  
  roc2 <- roc(model_data2$alc_binary, model_data2$pred)
  
  roc_list[[i]] <- roc2
  auc_values[i] <- auc(roc2)
  labels[i] <- paste0(model_labels[which(scores == score_name)], " + age+gender")
  i <- i + 1
}

# Grafic
pdf("ROC_summary_for_3_PRS.pdf", width = 7, height = 7)

#par(bg = "#F5F5F5")

plot(roc_list[[1]],
     col = colors[1],
     lwd = 2,
     main = "ROC curves for PRS models",
     legacy.axes = TRUE,
       xlab = "True positive rate",
      ylab = "False positive rate")

for (j in 2:length(roc_list)) {
  plot(roc_list[[j]],
       col = colors[j],
       lwd = 2,
       add = TRUE)
}

abline(a = 0, b = 1, lty = 2, col = "gray")

# --- AUC  ---
x_pos <- 0.2
y_pos <- seq(0.55, 0.35, length.out = length(auc_values))

for (k in 1:length(auc_values)) {
  text(x = x_pos,
       y = y_pos[k],
       labels = paste0("AUC: ", round(auc_values[k], 3)),
       col = colors[k],
       cex = 0.9,
       adj = 0)
}

# --- Legend (Model) ---
legend("bottomright",
       legend = labels,
       col = colors,
       lwd = 3,
       title = "Model",
       cex = 0.7,
       bty = "n")

dev.off()
```

The graph is plotted separately for three models without predictors and for three with predictors
```{r}
library(pROC)

model_labels <- c(
    "PGS000204",
    "PGS002752",
    "PGS003376"
)

scores <- c("PGS000204_hmPOS_GRCh37_SUM",
            "PGS002752_hmPOS_GRCh37_SUM",
            "PGS003376_hmPOS_GRCh37_SUM")

roc_prs <- list()
roc_cov <- list()

auc_prs <- c()
auc_cov <- c()

labels_prs <- c()
labels_cov <- c()

colors <- c("black", "red", "green")

i <- 1

for (score_name in scores) {
    
    merged_data[[paste0(score_name, "_z")]] <- scale(merged_data[[score_name]])
    
    ### PRS only
    model1 <- glm(
        as.formula(paste("alc_binary ~", paste0(score_name, "_z"))),
        family = binomial(),
        data = merged_data
    )
    
    model_data1 <- model1$model
    model_data1$pred <- predict(model1, type = "response")
    
    roc1 <- roc(model_data1$alc_binary, model_data1$pred)
    
    roc_prs[[i]] <- roc1
    auc_prs[i] <- auc(roc1)
    labels_prs[i] <- model_labels[which(scores == score_name)]
    
    
    ### PRS + covariates
    model2 <- glm(
        as.formula(paste("alc_binary ~", paste0(score_name, "_z"), "+ age + gender")),
        family = binomial(),
        data = merged_data
    )
    
    model_data2 <- model2$model
    model_data2$pred <- predict(model2, type = "response")
    
    roc2 <- roc(model_data2$alc_binary, model_data2$pred)
    
    roc_cov[[i]] <- roc2
    auc_cov[i] <- auc(roc2)
    labels_cov[i] <- paste0(model_labels[which(scores == score_name)], " + age+gender")
    
    i <- i + 1
}

########################################
# 🔹 1. PRS only
########################################

plot(roc_prs[[1]],
     col = colors[1],
     lwd = 2,
     main = "ROC curves",
     legacy.axes = TRUE,
     cex.main = 2,
     xlab = "False positive rate",
     ylab = "True positive rate",
     cex.lab = 1.5,
     cex.axis = 1.5)

for (j in 2:length(roc_prs)) {
    plot(roc_prs[[j]],
         col = colors[j],
         lwd = 2,
         add = TRUE)
}

#abline(a = 0, b = 1, lty = 2, col = "gray")

legend(x = 0.7, y = 0.4,
       legend = labels_prs,
       col = colors,
       lwd = 6,
       title = "Model",
       cex = 1.2,
       bty = "n")

########################################
# 🔹 2. PRS + covariates
########################################

plot(roc_cov[[1]],
     col = colors[1],
     lwd = 4,
     main = "ROC curves",
     legacy.axes = TRUE,
     cex.main = 2,
     xlab = "False positive rate",
     ylab = "True positive rate",
     cex.lab = 1.5,
     cex.axis = 1.5)

for (j in 2:length(roc_cov)) {
    plot(roc_cov[[j]],
         col = colors[j],
         lwd = 2,
         add = TRUE)
}

#abline(a = 0, b = 1, lty = 2, col = "gray")

legend(x = 0.8, y = 0.33,
       legend = labels_cov,
       col = colors,
       lwd = 6,
       title = "Model",
       cex = 1.0,
       bty = "n")
```

COMPARISON OF THE THREE BEST MODELS WITH ONE ANOTHER BY USING THE DELONG TEST
is used to compare two ROC curves, i.e., to check whether their AUCs differ statistically significantly.
Δ=AUC1−AUC2
SE(Δ)=Var(AUC1)+Var(AUC2)−2Cov(AUC1,AUC2)
Z=Δ\SE(Δ)Δ


The p-value is taken from a normal distribution
```{r}
library(pROC)

# Vectors for results
model_names <- labels
auc_vals <- auc_values

# We will calculate p-values only for pairs (PRS vs adjusted)
p_values <- rep(NA, length(roc_list))

for (i in seq(1, length(roc_list), by = 2)) {
  
  roc_prs <- roc_list[[i]]
  roc_adj <- roc_list[[i + 1]]
  
  test <- roc.test(roc_prs, roc_adj, method = "delong")
  
  p_values[i] <- test$p.value
  p_values[i + 1] <- test$p.value
}

results_table <- data.frame(
  Model = model_names,
  AUC = round(as.numeric(auc_vals), 3),
  DeLong_p_value = signif(p_values, 3)
)

results_table
```

