---
editor_options: 
  markdown: 
    wrap: 72
---

```{r,warning=FALSE, echo=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(bigsnpr)
library(bigstatsr)
```

```{r,echo=FALSE}
system("plink")
```

```{r}
head(fam <- fread('data/penncath.fam'))
```

Family ID Individual ID Paternal ID Maternal ID Sex (1=male; 2=female;
other=unknown) Phenotype

```{r}
head(clinical <- fread('data/penncath.csv'))
```

CAD=1 means that the subject has coronary artery disease), and some
covariates (age, triglycerides, HDL and LDL cholesterol levels

```{r}
bim <- fread('data/penncath.bim')
head(bim)
```

chromosome (1-22, X, Y or 0 if unplaced) rs# or snp identifier Genetic
distance (morgans) Base-pair position (bp units) Allele 1 (usually minor
allele, or ‘effect’ allele) Allele 2 (usually major allele) -9 - missed
data

```{r}
snp_readBed("data/penncath.bed")
penncath <- snp_attach("data/penncath.rds")

# check out this bigSNP object 
str(penncath)
```

create the .rds and .bk files

**Basic quality control steps**

1)Chromosome check

```{r}
penncath$map$chromosome |> table()
```

2)Missing data

```{r}
class(penncath$genotypes) 
```

```{r}
?bigstatsr::big_counts # handy function to summarize data 
```

```{r}
big_counts(penncath$genotypes, ind.row = 1:10, ind.col = 1:10) # в строках snp ( byrow = FALSE)
```

```{r}
snp_stats <- big_counts(penncath$genotypes)
```

```{r}
dim(snp_stats) 
boxplot(snp_stats[4,]) # распределение количества пропущенных генотипов по SNP
summary(snp_stats[4,]) 
```

exclude SNPs with \>10% missing data

```{r}
sample_stats <- big_counts(penncath$genotypes, byrow = TRUE) # byrow = TRUE-  она считает статистики по образцам (индивидам)
# cell [1,1] is showing the # of people with "0" genotype at the 1st SNP
sample_stats[, 1:10]

boxplot(sample_stats[4,])
summary(sample_stats[4,])
```

no outlier values

3)  Heterozygosity check

```{r}
allele_dat <- sweep(x = sample_stats[1:3,], 
                    # ^ leave off 4th row -- don't count NAs here
                    MARGIN = 1,
                    STATS = c(0, 1, 0),
                    FUN = "*") |> colSums()

boxplot(allele_dat/ncol(snp_stats))
```

```{r}
hist(allele_dat/ncol(snp_stats),
     main = "Zygosity",
     xlab = "Proportion of samples which are heterozygous") # should be bell-curved shaped
```

4)  Minor allele frequency (MAF) filtering

exclude SNPs that have low variation

```{r}
hist(snp_stats[1,]) #  в стобце 1 - гомозиготы
summary(snp_stats[1,])
```

5)  Hardy-Weinberg equilibrium filtering

------------------------------------------------------------------------

```{r}
system("plink --bfile data/penncath --make-bed --out data/penncath_clean")
```

```{r}
path_to_qc_penncath_bed <- snp_plinkQC(
  plink.path = "plink", # again, you may need to change this according to your machine!
  prefix.in = "data/penncath_clean", # input data
  prefix.out = "data/qc_penncath", # creates *new* rds with quality-controlled data
  maf = 0.01, # filter out SNPs with MAF < 0.01
  geno = 0.1, # filter out SNPs missing more than 10% of data
  mind = 0.1, # filter out SNPs missing more than 10% of data,
  hwe = 1e-10, # filter out SNPs that have p-vals below this threshold for HWE test
  autosome.only = TRUE # we want chromosomes 1-22 only
  
)
```

```{r}
snp_readBed("data/qc_penncath.bed")
```

**Imputation**

```{r}
obj <- snp_attach("data/qc_penncath.rds")
# double check dimensions 
dim(obj$fam)
# [1] 1401    6
dim(obj$map) 
# [1] 696644      6
dim(obj$genotypes)
```

```{r}
snp_stats <- bigstatsr::big_counts(obj$genotypes)
# ^ bigstatsr is loaded with bigsnpr, just being explicit here for didactic purposes
colnames(snp_stats) <- obj$map$marker.ID
snp_stats[,1:5] # shows # of 0, 1, 2, and NA values for 1st 5 SNPs

(any_missing <- sum(snp_stats[4,] != 0)) # shows # of SNPs with NA values 
# у 565138 все еще есть пропущенные значения (хотя и менее 10%)
call_rates <- colSums(snp_stats[1:3,])/colSums(snp_stats) 
head(call_rates)
```

```{r}
# impute based on mode
obj$geno_imputed <- snp_fastImputeSimple(Gna = obj$genotypes,
                                          method = "mode",
                                          ncores = nb_cores())

# save imputed values 
obj$geno_imputed$code256 <- bigsnpr::CODE_IMPUTE_PRED

# look to see that imputation was successful:
imp_snp_stats <- big_counts(X.code = obj$geno_imputed)
imp_snp_stats[,1:5] # all 0s in NA row 
```

```{r}
obj <- bigsnpr::snp_save(obj)
```

**Principal component analysis (PCA)**

```{r}

svd_X <- big_SVD(obj$geno_imputed, # must have imputed data
                 big_scale(), # centers and scales data -- REALLY IMPORTANT! 
                 k = 10 # use 10 PCs for now -- can ask for more if needed
                 )

saveRDS(svd_X, "data/svd_X.rds")
```

```{r}
svd_X <- readRDS("data/svd_X.rds")

str(svd_X)

# scree plot
plot(svd_X) 
```

```{r}
plot(svd_X, type = "scores")
```

```{r}
clinical <- fread('data/penncath.csv') |>
  mutate(across(.cols = c(sex, CAD),
                .fns = as.factor))
dplyr::glimpse(clinical)


plot(svd_X, type = "scores") +
  aes(color = clinical$sex) +
  labs(color = "Sex")
```

Family relatedness

```{r}
K <- big_tcrossprodSelf(X = obj$geno_imputed,
                        fun.scaling = big_scale())

save(K, file = "data/K.RData")
```

```{r}
library(corrplot)
corrplot::corrplot(K[1:50, 1:50]*(1/ncol(obj$geno_imputed)),
                   is.corr = FALSE,
                   tl.pos = "n")
```

**SNP testing (marginal approach)**

```{r}
obj$fam <- dplyr::full_join(x = obj$fam,
                            y = clinical,
                            by = c('family.ID' = 'FamID'))
```

```{r}
# load principal components 
svd_X <- readRDS(file = "data/svd_X.rds")

# calculate the PCs                          
pc <- sweep(svd_X$u, 2, svd_X$d, "*")
dim(pc) # will have same number of rows as U
# [1] 1401   10
names(pc) <- paste0("PC", 1:ncol(pc))
```

Implement tests

```{r}
# Fit a logistic model between the phenotype and each SNP separately
# while adding PCs as covariates to each model
obj$fam$CAD <- as.numeric(as.character(obj$fam$CAD))
obj.gwas <- big_univLogReg(X = obj$geno_imputed,
                           y01.train = obj$fam$CAD,
                           covar.train = pc[,1:5],
                           ncores = nb_cores())

# this takes a min or two, so I will save these results 
saveRDS(object = obj.gwas, file = "data/gwas.rds")
```

```{r}
obj.gwas <- readRDS("data/gwas.rds")
```

Visualizing

QQ plot

```{r}
qq <- bigsnpr::snp_qq(obj.gwas)
(qq)
```

```{r}
viridis22 <- c(rep(c("#fde725", "#90d743", "#35b779", "#21918c", "#31688e", 
                "#443983", "#440154"), 3), "#fde725")

manh <- snp_manhattan(gwas = obj.gwas,
              infos.chr = obj$map$chromosome,
              infos.pos = obj$map$physical.pos,
              colors = viridis22)

(manh)
```

```{r}
signif_threshold <- 5e-8 
suggest_threshold <- 5e-6 

(manh + 
  geom_hline(yintercept = -log10(signif_threshold),
             # note: plot y-axis is on -log10 scale 
             color = "#35b779",
             lty = 2) + 
  geom_hline(yintercept = -log10(suggest_threshold),
             color = "#443983",
             lty = 2)
)
```
```{r}
# Подключаем нужные библиотеки
library(ggplot2)
library(data.table)

# Загружаем данные
gwas <- fread("alcohol_sub_gwas.ss.gz")

# Преобразуем P-значения в -log10(P) для оси Y
gwas$negLogP <- -log10(gwas$P)

# Построим Manhattan plot с порогами значимости
ggplot(gwas, aes(x = POS, y = negLogP, color = as.factor(CHROM))) +
  geom_point(alpha = 0.6, size = 0.5) + 
  scale_color_manual(values = rep(c("blue", "orange"), length.out = length(unique(gwas$CHROM)))) +  # Чередование цветов для хромосом
  geom_hline(yintercept = -log10(5e-8), color = "red", linetype = "dashed") +  # Линия для порога значимости P-value ≤ 5e-8
  geom_hline(yintercept = -log10(5e-5), color = "blue", linetype = "dashed") + # Линия для порога P-value ≤ 5e-5
  facet_wrap(~ CHROM, scales = "free_x", ncol = 5) +  # Разбиваем по хромосомам
  theme_minimal() + 
  labs(
    title = "Manhattan Plot for AUD",
    x = "Chromosome Position",
    y = "-log10(P-value)"
  ) + 
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text.x = element_text(size = 8)
  )

```

